---
title: "R Notebook"
output: html_notebook
---

# Load libraries 

```{r,message=FALSE,warning=FALSE}

library(tidyverse)
library(tidymodels)
library(janitor)
library(skimr)
library(kableExtra)
library(GGally)
library(kableExtra) 
library(vip)        
library(fastshap)   
library(MASS)
library(ISLR)
library(tree)
library(dplyr)
library(ggplot2)
library(factoextra)
library(ggpubr)
```

# Read Data

```{r,message=FALSE,warning=FALSE}
donation = read_csv("DonorMerge_Final.csv") %>%
  clean_names()
donor = read.csv('Donations.csv') %>%
  clean_names()
donor = donor %>%
  inner_join(donation%>%dplyr::select(projectid|is_exciting),by='projectid')
skim(donor)
skim(donation)
```

# Data Preparation

```{r,message=FALSE,warning=FALSE}
#variable removing
donor_del = c('donor_zip','donationid','projectid','donor_acctid','donor_city','donation_message','donation_timestamp')
donation_del = c('school_ncesid','projectid','teacher_acctid','schoolid','school_city','school_district','school_county','school_zip')
donation = donation %>%
  dplyr::select(!donation_del)
donor = donor %>%
  dplyr::select(!donor_del)

# data transformation
donation = donation %>%
  mutate(date_posted = as.Date(date_posted,format='%m/%d/%Y')) %>%
  mutate(days = as.numeric(word(round(difftime('2022-12-02',date_posted,units="days"),0)))) %>%
  dplyr::select(!date_posted) 

```

#Exploring variables

## Exploring target
```{r,message=FALSE,warning=FALSE}
summary <-  donation %>%
  count(is_exciting) %>%
  mutate(pct = n/sum(n))
summary
summary %>%
  ggplot(aes(x=factor(is_exciting),y=pct)) +
  geom_col()  + 
  geom_text(aes(label = round(pct,2)) , vjust = 1, colour = "white") + 
  labs(title="Exciting Program", x="Is Exciting", y="PCT")+theme_bw() + theme(panel.grid=element_blank())+ scale_fill_brewer(palette = 5)
```
## Donation exploring
### Exploring numeric 
```{r}
num_explore <- function(col){
  donation %>%
    ggplot(., aes(x=!!as.name(col), y=factor(is_exciting))) + geom_boxplot()+theme_bw() + theme(panel.grid=element_blank())
}

# numeric_char=c('customer_service_calls','late_payments','number_phones')

for (column in names(donation %>% select_if (is.numeric))){
  if (column != 'is_exciting'){
    num <- num_explore(column)
    print(num)
  }
}

# checking outliers

donation %>%
  filter(total_price_excluding_optional_s<1000) %>%
  ggplot(., aes(x=total_price_excluding_optional_s, y=factor(is_exciting))) + 
  geom_boxplot()

donation %>%
  filter(total_price_including_optional_s<1000) %>%
  ggplot(., aes(x=total_price_including_optional_s, y=factor(is_exciting))) + 
  geom_boxplot()

donation %>%
  filter(students_reached<1000) %>%
  ggplot(., aes(x=students_reached, y=factor(is_exciting))) + 
  geom_boxplot()

# removing outliers
donation = donation %>%
  filter(total_price_excluding_optional_s<1000) %>%
  filter(total_price_including_optional_s<1000) %>%
  filter(students_reached<1000)
```

### Explore character & logical

```{r, warning=FALSE, message=FALSE}

char_explore <- function(col){
  donation %>%
    ggplot(., aes(!!as.name(col))) + 
    geom_bar(aes(fill = factor(is_exciting)), position = "fill") +theme_bw() + theme(panel.grid=element_blank()) + scale_fill_brewer(palette = 5) +
  coord_flip()
}

#character var
for (column in names(donation %>% select_if(is_character))){
    chrt <- char_explore(column)
    print(chrt)
}



```
## Donor exploring
### Exploring numeric 
```{r}


# checking outliers
donor%>%
  filter(donation_to_project<200) %>%
  ggplot(., aes(x=donation_to_project))+ geom_boxplot()

# 1,048,575
# 1,048,558 <20000
# 1,048,535 <10000
# 1,048,381 <5000
# 1,045,399 <1000
# 1,031,794 <500
# 966,876 <200

donor%>%
  #filter(donation_optional_support<50) %>%
  ggplot(., aes(x=donation_optional_support))+ geom_boxplot()

# 1,046,169 <200
# 1,035,724 <100
# 999,629 <50

donor%>%
  filter(donation_total<250) %>%
  ggplot(., aes(x=donation_total))+ geom_boxplot()

# 949,651 <200
# 973,011 <250

# removing outliers
donor = donor %>%
  filter(donation_to_project<200) %>%
  filter(donation_optional_support<50) %>%
  filter(donation_total<250)
```

### Explore character & logical

```{r, warning=FALSE, message=FALSE}

char_explore <- function(col){
  donor %>%
    ggplot(., aes(!!as.name(col))) + geom_bar()+theme_bw() + 
    theme(panel.grid=element_blank()) + 
    scale_fill_brewer(palette = 5)
}

for (column in names(donor %>% select_if(is_character))){
    chrt <- char_explore(column)
    print(chrt)
}

```
# Clustering
## data processing

```{r}
recipe =recipe(is_exciting ~ .,
                     data=donor) %>%
  step_impute_median(all_numeric_predictors()) %>%
  step_impute_mode(all_nominal_predictors()) %>%
  step_zv(all_numeric_predictors()) %>%
  step_scale(all_numeric_predictors()) %>%
  step_dummy(all_nominal_predictors())

donor_bake = bake(recipe %>% prep(), new_data = donor,composition = "tibble")
skim(donor_bake)
```
## visually choose number of clusters

```{r}

set.seed(123)

# -- performs our train / test split 
split_fig <- initial_split(donor_bake, prop = 0.02)

# -- extract the training data form our banana split 
donor_fig <- training(split_fig)


# how many clusters

fviz_nbclust(donor_fig, kmeans, method="wss")

```

## build clusters

```{r}
set.seed(17)

clusters5 <- kmeans(donor_bake, 5, iter.max = 200, nstart = 5)
print(clusters5)


# visualize clusters

fviz_cluster(clusters5,donor_bake,ellipse.type="norm",geom="point")

```

## cluster detail
```{r}
cluster <- as.factor(clusters5$cluster)
clusters5
```

```{r}
tree.clusters=tree(cluster~., donor_bake)

summary(tree.clusters)

plot(tree.clusters)
text(tree.clusters,pretty=0)

```



# Exploring cluster
## explore cluster target
```{r}
ggplot(donor,aes(cluster))+geom_bar()
```
## explore cluster features
```{r}
# categorical
ggplot(donor,aes(is_teacher_acct))+geom_bar()
ggplot(donor,aes(is_teacher_acct))+geom_bar()+facet_wrap(~clusters5$cluster) 

###
ggplot(donor,aes(dollar_amount))+geom_bar()
ggplot(donor,aes(dollar_amount))+geom_bar()+facet_wrap(~clusters5$cluster) 

ggplot(donor,aes(payment_method))+geom_bar()
ggplot(donor,aes(payment_method))+geom_bar()+facet_wrap(~clusters5$cluster)  +
  theme(axis.text.x=element_text(angle=90))

ggplot(donor,aes(payment_included_campaign_gift_card))+geom_bar()
ggplot(donor,aes(payment_included_campaign_gift_card))+geom_bar()+facet_wrap(~clusters5$cluster) 

ggplot(donor,aes(via_giving_page))+geom_bar()
ggplot(donor,aes(via_giving_page))+geom_bar()+facet_wrap(~clusters5$cluster) 


# numeric
ggplot(donor,aes(x=donation_to_project))+geom_histogram(binwidth=10)
ggplot(donor,aes(x=donation_to_project))+geom_histogram(binwidth=10) + facet_wrap(~clusters5$cluster)

ggplot(donor,aes(x=donation_optional_support))+geom_histogram(binwidth=10)
ggplot(donor,aes(x=donation_optional_support))+geom_histogram(binwidth=10) + facet_wrap(~clusters5$cluster)

ggplot(donor,aes(x=donation_total))+geom_histogram(binwidth=10)
ggplot(donor,aes(x=donation_total))+geom_histogram(binwidth=10) + facet_wrap(~clusters5$cluster)


```


# LR
## Partition data

```{r}
set.seed(123)
donation_lr = donation %>%
  mutate_if(is.character, factor) %>%
  mutate_if(is.logical,factor) %>%
  mutate(is_exciting  = if_else(is_exciting=='TRUE',1,0)) %>%
  mutate(is_exciting = factor(is_exciting)) 
# -- performs our train / test split 
split <- initial_split(donation_lr, prop = 0.7)

# -- extract the training data form our banana split 
LR_train <- training(split)
# -- extract the test data 
LR_test <- testing(split)

sprintf("Train PCT : %1.2f%%", nrow(LR_train)/ nrow(donation_lr) * 100)

```

## Define Recipe 

```{r}
LR_recipe <- recipe(is_exciting ~ .,
                     data=LR_train) %>%
  step_impute_median(all_numeric_predictors()) %>%
  step_impute_mode(all_nominal(), -all_outcomes()) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_other(all_nominal_predictors()) %>%
  step_zv(all_predictors())
                                    
# eyeball recipe results 
LR_train_bake = bake(LR_recipe %>% prep(), new_data = LR_train)
LR_test_bake = bake(LR_recipe %>% prep(), new_data = LR_test)
```

## Fitting models
```{r}
logistic_glm <-logistic_reg(mode = "classification") %>%
                  set_engine("glm") %>%
                  fit(is_exciting ~ ., data=LR_train_bake)
tidy(logistic_glm) %>%
  mutate_at(c("estimate", "std.error", "statistic", "p.value"),round, 4) %>%
  filter(p.value < 0.01) 
```
## Evaluating model

```{r}
# score training
predict(logistic_glm , LR_train_bake, type = "prob") %>%
  bind_cols(predict(logistic_glm , LR_train_bake, type = "class")) %>%
  mutate(part = "train") %>%
  bind_cols(., LR_train_bake) -> scored_train

# -- score testing
predict(logistic_glm , LR_test_bake, type = "prob") %>%
  bind_cols(predict(logistic_glm ,  LR_test_bake, type = "class")) %>%
  mutate(part = "testing") %>%
  bind_cols(., LR_test_bake) -> scored_test

## Metrics (AUC / Accuracy / Log Loss)
bind_rows (scored_train, scored_test)  %>%
  group_by(part) %>%
  metrics(is_exciting, .pred_0, estimate = .pred_class) %>%
  filter(.metric %in% c('accuracy', 'roc_auc', 'mn_log_loss')) %>%
  pivot_wider(names_from = .metric, values_from = .estimate)

```

## Define Recipe for reduced model

```{r}


LR_steprecipe <- recipe(is_exciting ~ school_latitude	+			
one_non_teacher_referred_donor_g				+
great_messages_proportion				+
teacher_referred_count				+
non_teacher_referred_count	+			
school_charter	+			
teacher_teach_for_america	+			
fulfillment_labor_materials		+		
total_price_excluding_optional_s	+			
total_price_including_optional_s +
eligible_double_your_impact_matc		+		
eligible_almost_home_match			+	
days	+
school_state +
school_metro+
resource_type,
                     data=LR_train)%>%
  step_impute_mode(all_nominal(), -all_outcomes()) %>%
  step_impute_median(all_numeric()) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  prep()

# eyeball recipe results 
LR_train_bake_step = bake(LR_steprecipe %>% prep(), new_data = LR_train)
LR_test_bake_step = bake(LR_steprecipe %>% prep(), new_data = LR_test)

LR_steprecipe
```

## Fitting logistic regression

```{r,warning=FALSE, message=FALSE}

logistic_glm_step <-logistic_reg(mode = "classification") %>%
                  set_engine("glm") %>%
                  fit(is_exciting ~ ., data=LR_train_bake_step)

tidy(logistic_glm_step) %>%
  mutate_at(c("estimate", "std.error", "statistic", "p.value"),round, 4) 

logistic_glm_step %>%
  vi()
```
## Evaluating model

```{r}
# score training
predict(logistic_glm_step, LR_train_bake_step, type = "prob") %>%
  bind_cols(predict(logistic_glm_step, LR_train_bake_step, type = "class")) %>%
  mutate(part = "train") %>%
  bind_cols(., LR_train_bake_step) -> scored_train_step

# score testing
predict(logistic_glm_step, LR_test_bake_step, type = "prob") %>%
  bind_cols(predict(logistic_glm_step,  LR_test_bake_step, type = "class")) %>%
  mutate(part = "testing") %>%
  bind_cols(., LR_test_bake_step) -> scored_test_step

## Metrics (AUC / Accuracy / Log Loss)
bind_rows (scored_train_step, scored_test_step)  %>%
  group_by(part) %>%
  metrics(is_exciting, .pred_0, estimate = .pred_class) %>%
  filter(.metric %in% c('accuracy', 'roc_auc', 'mn_log_loss')) %>%
  pivot_wider(names_from = .metric, values_from = .estimate)
```

# Random Forest
## Partition our data 70/30 
```{r}
# Save the split information for an 70/30 split of the data
bsplit <- initial_split(donation_lr, prop = 0.7)
train <- training(bsplit) 

test  <-  testing(bsplit)

```

## Recipe

```{r}
# write out the formula 


donation_recipe <-
  recipe(is_exciting ~ school_latitude	+			
one_non_teacher_referred_donor_g				+
great_messages_proportion				+
teacher_referred_count				+
non_teacher_referred_count	+			
school_charter	+			
teacher_teach_for_america	+			
fulfillment_labor_materials		+		
total_price_excluding_optional_s	+			
total_price_including_optional_s +
eligible_double_your_impact_matc		+		
eligible_almost_home_match			+	
days	+
school_state +
school_metro+
resource_type,
                     data=train) %>%
  step_impute_median(all_numeric_predictors()) %>% # missing values numeric 
  step_novel(all_nominal_predictors()) %>% # new factor levels 
  step_unknown(all_nominal_predictors()) %>% # missing values 
  step_dummy(all_nominal_predictors(), one_hot = TRUE) %>%
  step_nzv(all_predictors()) 

## Check the recipe results m
bake(donation_recipe %>% prep(),train %>% sample_n(1000))

```

## Model & Workflow

```{r}

rf_model <- rand_forest(trees = 100, min_n = 20) %>%
   set_mode("classification") %>%
   set_engine("ranger", 
              num.threads = 8, 
              max.depth = 10, 
              importance="permutation")

rf_workflow <- workflow() %>%
  add_recipe(donation_recipe) %>%
  add_model(rf_model) %>%
  fit(train)

rf_workflow %>%
  extract_fit_parsnip() %>%
  vi()

```

## Evaluating model
### Accuracy & ROC
```{r}
# score training
predict(rf_workflow, train, type = "prob") %>%
  bind_cols(predict(rf_workflow, train, type = "class")) %>%
  mutate(part = "train") %>%
  bind_cols(., train) -> scored_train_step

# score testing
predict(rf_workflow, test, type = "prob") %>%
  bind_cols(predict(rf_workflow,  test, type = "class")) %>%
  mutate(part = "testing") %>%
  bind_cols(., test) -> scored_test_step

## Metrics (AUC / Accuracy / Log Loss)
bind_rows (scored_train_step, scored_test_step)  %>%
  group_by(part) %>%
  metrics(is_exciting, .pred_0, estimate = .pred_class) %>%
  filter(.metric %in% c('accuracy', 'roc_auc', 'mn_log_loss')) %>%
  pivot_wider(names_from = .metric, values_from = .estimate)
```

### Precision & Recall

```{r}
scored_train_step %>%
  precision(is_exciting, .pred_class) %>%
  mutate(part="training") %>%
  bind_rows(
  scored_test_step %>%
  precision(is_exciting, .pred_class) %>%
    mutate(part="testing") 
  )

scored_train_step %>%
  recall(is_exciting, .pred_class) %>%
  mutate(part="training") %>%
  bind_rows(
  scored_test_step %>%
  recall(is_exciting, .pred_class) %>%
    mutate(part="testing") 
  )
```

